Inferring the gravitational potential of the Milky Way with a few precisely measured stars

### 2014-03-15 text/paper2/ms.tex
added more notes on experiments

Figures~\ref{fig:}-\ref{fig:} show the posterior distributions for these five parameters for each of the the four mass progenitors.

\subsection{\spitzer + \gaia}
We now take the same eight stars used in the previous experiment and ``observe'' them with optimistic observational uncertainties. We take these stars to be RR Lyrae variables which are known to be excellent distance indicators using the mid-infrared period-luminosity relation \citep[as shown in, e.g.,][]{madore12}; $\sim$100 RR Lyrae associated with the Sgr stream and $\sim$30 associated with the Orphan stream will be observed with \spitzer as part of the SMASH survey \citep{smashprop} with expected fractional distance uncertainties around $\sim$2\%. These stars will also be included in the \gaia\, proper motion catalog: at a distance of $\sim$50~kpc, a typical RR Lyrae will have a tangential velocity error around $\sim$20~km/s.

For this experiment, we (1) assume the stars are bright distance indicators --- RR Lyrae --- such that the fractional distance uncertainty is $2\%$; (2) neglect the uncertainty in angular position (for a typical RR Lyrae at 50~kpc this is $\sim$$10^{-7}$~deg); (3) assume we can measure radial velocities to these stars with 5~km/s uncertainty; and (4) compute the sky-averaged \gaia\, proper-motion uncertainty for each star assuming an F0V spectral type using the \texttt{PyGaia}\footnote{https://github.com/agabrown/PyGaia} code and use this uncertainty for both components of proper motion. % TODO: acknowledge Gaia science team, Anthony Brown, Gaia data challenge?
The observed positions and velocities of these stars are shown in Figure~\ref{fig:} (red, compare to the unobserved black points).

@@

\acknowledgements
Aspen?\\
APW: NSF GRFP, Sigma Xi\\
Columbia HPC\\

>>>>

Figures~\ref{fig:}-\ref{fig:} show the posterior distributions for these five parameters for each of the the four mass progenitors...

\subsection{\spitzer + \gaia}
For this and subsequent experiments, we use only the $m=2.5\times10^8~M_\odot$ simulation. We take the same eight stars used in the previous experiment and ``observe'' them with optimistic observational uncertainties. We take these stars to be RR Lyrae variables which are known to be excellent distance indicators via the mid-infrared period-luminosity relation \citep[as shown in, e.g.,][]{madore12}. $\sim$100 RR Lyrae associated with the Sgr stream and $\sim$30 associated with the Orphan stream will be observed with \spitzer as part of the SMASH survey \citep{smashprop} with expected fractional distance uncertainties around $\sim$2\%. These stars will also be included in the \gaia\, proper motion catalog: at a distance of $\sim$50~kpc, a typical RR Lyrae will have a tangential velocity error around $\sim$20~km/s.

For this experiment, we (1) assume the stars are bright distance indicators --- RR Lyrae --- such that the fractional distance uncertainty is $2\%$; (2) neglect the uncertainty in angular position (for a typical RR Lyrae at 50~kpc this is $\sim$$10^{-7}$~deg for \gaia); (3) assume we can measure radial velocities to these stars with 5~km/s uncertainty; and (4) compute the sky-averaged \gaia\, proper-motion uncertainty for each star assuming an F0V spectral type using the \texttt{PyGaia}\footnote{https://github.com/agabrown/PyGaia} code and use this uncertainty for both components of proper motion. % TODO: acknowledge Gaia science team, Anthony Brown, Gaia data challenge?
The observed positions and velocities of these stars are shown in Figure~\ref{fig:} (red, compare to the unobserved black points). We further assume that we know the tail assignment for each star, $\beta$. We observe the position of the progenitor with the same observational uncertainties, though in reality some coordinates will have higher or lower uncertainties.

This experiment has the four potential parameters, the Lagrange point offset, four phase-space coordinate parameters for each star (32 total), and four phase-space coordinate parameters for the progenitor --- 41 parameters in total.

@@

\acknowledgements
Aspen?\\
Gaia data challenge / Surrey\\
APW: NSF GRFP, Sigma Xi\\
Columbia HPC\\
